---
title: "Literate Programming"
author: "Niklas Wulms"
date: "6 May 2019"
bibliography: Rpackages.bib
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float:
      toc_collapsed: yes
    code_folding: hide


---

<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):after {
  content: '';
}
</style>


```{r eval=FALSE, echo=FALSE}
# add here your libraries - they are automatically added to "Rpackages.bib" file and can be referred to!

write.bib(c("rmarkdown", "bibtex", "htmlwidgets", "stargazer", "DT", "knitr", "summarytools", "ggplot2", "skimr", "plotly", "tidyverse"))

```


# RMarkdown basics

## Your task:

* install R
* install RStudio
* Click File
    + New File
    + RMarkdown [@rmarkdown1, rmarkdown2]...
    + install libraries/dependencies
* open the .Rmd (RMarkdown) file of this HTML 
    + run the code chunk with `install.packages()` manually
    + klick on __Knit__ [@knitr1, @knitr2, @knitr3]
    + compare .Rmd with .html output
    + understand, what is happening
    + copy/paste and use provided functions
* create your first RMarkdown document
* download the RStudio RMarkdown cheat sheet




## RMarkdown Example: "Tools for regression analysis in R"

If you opened the ".Rmd" file in RStudio, you can run the code by clicking the "play" button in the code chunks. If you open the html file, you can copy & paste the code and interact with some widgets, but are not able to change anything.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. 

You can add an __R code chunk__ by keyboard shortcuts (Win: Control + Alt + I, macOS: Cmd + Option + I) like this. 

## Chunk options

We entered the code `print("Hello world!")` and it is evaluated.

```{r Visible Chunk}

print("Hello world!")
```

__Invisible chunk:__ Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing (_echo_) of the code chunk. Commands are executed and output is generated. 

```{r Invisible Chunk, echo=FALSE}
print("You are not able to see the code chunk producing this output in the html!")
```

__Inactive chunk:__ Note that the `eval = FALSE` parameter was added to the code chunk to prevent _evaluation_ of the R code that generated the output. Commands are not executed, no output is generated.

```{r Inactive chunk, eval=FALSE}
print("This command is not evaluated. But you are able to see the chunk.")
```






# Preparation {.tabset}

The used libraries can be installed and are loaded here.

## Install Libraries

`eval=FALSE` in chunk turns code evaluation during __"Knit"__ off!!! You don't need to install the packages during each __Knit__ again and again.

__BUT__: To "Knit" the .Rmd you need to run this chunk manually to install the packages.

```{r eval=FALSE}
install.packages("glue")
install.packages("colorspace")
install.packages("skimr")
install.packages("summarytools")
install.packages("tidyverse")
install.packages("DT")
install.packages("stargazer")
install.packages("plotly")
```

## Libraries

```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(skimr)
library(pander)
library(summarytools)
library(tidyverse)
library(DT)
library(stargazer)
library(plotly)
```

## Citation of used libraries

```{r}
citation()
citation("skimr")
citation("summarytools")
citation("tidyverse")
citation("DT")
citation("stargazer")
citation("plotly")
```



# Dataset selection

## Datasets 

Here you can find example datasets for demonstrating functions.

In reality you would:  
 
- load or import your data  
- clean and tidy up your data  
- calculate variables  
- check variable types  

## Find example datasets

```{r explore possible datasets, eval=FALSE}
# to look at R-base datasets
data()
# to look at all datasets included in your instance of R throug additional installed packages
data(package = .packages(all.available = TRUE))
```


I selected the __ChickWeight__ dataset for further demonstration. You can take your own datasets and explore them with the functions mentioned here.

```{r}
?ChickWeight
```


# Data in the ChickWeight dataset 


## static tables {.tabset}

### Base-R

```{r}
head(ChickWeight, n = 20)
```

### package "tibble" (tidyverse [@tidyverse])

```{r}
glimpse(ChickWeight)
```

## interactive tables {.tabset}

### RMarkdown "paged_table"

Default setting if `df_print: paged` is added to the header!

```{r}
rmarkdown::paged_table(ChickWeight)
```

### package "DT" (Datatable [@DT])

```{r}
datatable(ChickWeight)
```

### customized DT [@DT]

```{r}
ChickWeight %>%
    datatable(filter = 'top', 
            extensions = list('Buttons' = NULL,
                              'Scroller' = NULL),
            
            options = list(
                           dom = 'Bfrtip', 
                           buttons = list(c('copy', 'csv', 'excel', 'pdf', 'print'), 
                                          I('colvis')),
                           deferRender = TRUE,
                           scrollX = 400,
                           scrollY = 600, 
                           scroller = TRUE,
                           pageLength = 10
                           )
    )
```




# Summary statistics {.tabset}

## Base-R

```{r Summary}
summary(ChickWeight)
```

## Package "skimr" [@skimr]

```{r Skimr function, results = "asis", warning=FALSE}
panderOptions('table.split.table', Inf)

 skim(ChickWeight) %>%
   pander(table.split.table = Inf, table.split.cells = Inf)
```

## Package "Stargazer" [@stargazer]

```{r, results="asis"}
stargazer(ChickWeight, type = "html")
```



## Package "summarytools" [@summarytools]

```{r, results="asis"}
ChickWeight %>%
  dfSummary(style = "grid", plain.ascii = FALSE, graph.magnif = 0.75, tmp.img.dir = "/tmp")


```



# Plots using ggplot2 [@ggplot2]

## Static plots {.tabset}

### Lineplot

weight ~ time for each chicken (`geom_line()`)

```{r}
ChickWeight %>%
  ggplot(aes(x = Time, y = weight, col = Chick)) +
  geom_line() #+
#  geom_smooth(aes(group = Diet)) +
#  facet_grid( ~ Diet) +
#  theme(legend.title = element_blank(),
#        legend.position = "none")
```


### remove legend

```{r}
ChickWeight %>%
  ggplot(aes(x = Time, y = weight, col = Chick)) +
  geom_line() +
#  geom_smooth(aes(group = Diet)) +
#  facet_grid( ~ Diet) +
  theme(legend.title = element_blank(),
        legend.position = "none")
```


### include fits

`loess fit` (`geom_smooth()`)

```{r}
ChickWeight %>%
  ggplot(aes(x = Time, y = weight, col = Diet)) +
#  geom_line(aes(x = Time, y = weight, col = Chick)) +
  geom_smooth() +
#  facet_grid( ~ Diet) +
  theme(legend.title = element_blank(),
        legend.position = "none")


ChickWeight %>%
  ggplot(aes(x = Time, y = weight, col = Diet)) +
#  geom_line(aes(x = Time, y = weight, col = Chick)) +
  geom_smooth(method = "lm") +
#  facet_grid( ~ Diet) +
  theme(legend.title = element_blank(),
        legend.position = "none")
```


### create subplots

split the data by `diet` (`facet_grid(+ Diet)`)

```{r}

plot1 <- ChickWeight %>%
  ggplot(aes(x = Time, y = weight, col = Chick)) +
  geom_line() +
  geom_smooth(aes(group = Diet)) +
  facet_grid( ~ Diet) +
  theme(legend.title = element_blank(),
        legend.position = "none")

plot1

```

### Boxplots

```{r}


plot2 <- ChickWeight %>%
  ggplot(aes(x = Time, y = weight, group = Time)) +
  geom_boxplot() +
#  geom_smooth(aes(group = Diet)) +
  facet_grid( ~ Diet) +
  theme(legend.title = element_blank(),
        legend.position = "none")


plot2

```

## interactive plots with plotly [@plotly] {.tabset}


### subplots 

```{r}
ggplotly(plot1)


```

### boxplots 
```{r}
ggplotly(plot2)
```


# Fitting a linear / multiple regression model {.tabset}

## linear regression model

`lm(formula = dependent_variable ~ predictor/observed_variable,
data = your_data)`

```{r}
linear_regression_model <- lm(formula = weight ~ Time,
                              data = ChickWeight)

summary(linear_regression_model)


```

## plot diagnostics

```{r}
par(mfrow = c(2,2))
plot(linear_regression_model)
```


## multiple regression model

add with `+ var` further variables to the model

`lm(formula = dependent_variable ~ 1st_var + 2nd_var + more_variables, data = your_data)`



```{r}
multiple_regression_model <- lm(formula = weight ~ Time + Diet,
                                data = ChickWeight)

summary(multiple_regression_model)
```

## plot diagnostics

```{r}
par(mfrow = c(2,2))
plot(multiple_regression_model)
```

## Stargazer Publication Ready Tables [@stargazer]

```{r results = 'asis'}
stargazer(linear_regression_model, multiple_regression_model,
          title = "Linear and Multiple Regression - ChickWeight",
          column.labels = c("Linear Model", "Multiple Model"),
          type = "html",
          ci = TRUE,
          digit.separator = ",",
          report = c("vcs*"), 
          single.row = TRUE
          )
```

# Questions / Feedback


## References